---
title: "ManyBabies1 Participating Labs Map"
author: "Meghan Mastroberardino, with the help of Esther Schott and Krista Byers-Heinlein: Concordia Infant Research Lab"
output: html_notebook
---

This is a map of labs countries and cities that have participated in ManyBabies 1

#cleanworkspace and load required packages
```{r}
rm(list=ls()) # Clear workspace
library(tidyverse)
library(rworldmap)
library(ggmap)
```

#importing dataset
```{r}
#importing participating las spreadsheet
wide_dataset_labs <- read.delim("../data/ManyBabiesQuestionnaireSpreadsheetAuthoritativeMay22-2018cleaned.csv",sep = ",") # Laboratory questionnaire - hand clean by Melanie

cities_dataframe <- read.delim("../data/ManyBabies_participatinglabs_cities_dataframe.csv", sep = ",") # Dataframe linking University names to their cities - must be updated by hand

```

# function to get country from longitude and latitude coordinates
```{r}
# Code from: https://stackoverflow.com/questions/41105042/using-coords2country-function-in-r-on-exclusive-economic-zones-not-country-bound

# The single argument to this function, points, is a data.frame in which:
#   - column 1 contains the longitude in degrees
#   - column 2 contains the latitude in degrees
coords2country = function(points)
{  
  countriesSP <- getMap(resolution='low')
  #countriesSP <- getMap(resolution='high') #you could use high res map from rworldxtra if you were concerned about detail

  # convert our list of points to a SpatialPoints object

  # pointsSP = SpatialPoints(points, proj4string=CRS(" +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"))

  #setting CRS directly to that from rworldmap
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  


  # use 'over' to get indices of the Polygons object containing each point 
  indices = over(pointsSP, countriesSP)

  # return the ADMIN names of each country
  indices$ADMIN  
  #indices$ISO3 # returns the ISO3 code 
  #indices$continent   # returns the continent (6 continent model)
  #indices$REGION   # returns the continent (7 continent model)
}
```


# looking up country info and cleaning dataset
```{r, , echo = FALSE}

clean_wide_dataset_labs <- wide_dataset_labs %>%
  select(labname, University, CityorCountry, Longitude, Latitude)

clean_wide_dataset_labs$country <- as.character(coords2country(clean_wide_dataset_labs[, c("Longitude", "Latitude")])) 

clean_wide_dataset_labs <- clean_wide_dataset_labs %>%
  mutate(country = recode(country, "United Kingdom" = "UK", "United States of America" = "USA")) #World code naming of countries and lab dataset names did not match

clean_wide_dataset_labs$country[clean_wide_dataset_labs$University =="University of Liverpool"] ="UK" #Lab forgot to put in country


```

# merging with cities dataframe
```{r}

cities_and_lat_log <- clean_wide_dataset_labs %>%
  full_join(cities_dataframe, by = c("University")) %>%
  distinct(labname, .keep_all = TRUE) %>% #get rid of labs repeated twice
  group_by(City) %>%
  mutate(n_in_city = as.integer(length(City))) %>% # Get number of labs/city
  arrange(n_in_city) %>%
  distinct(City, .keep_all = TRUE)
```

# count of frequency of countries participating
```{r}
MB.Nlabs = clean_wide_dataset_labs %>%
  group_by(country) %>% 
 summarise(Nlabs= n())
  
```

# count of frequency of cities particpating
```{r}
MBlabs.Nlabs_cities = cities_dataframe %>%
  group_by(City) %>%
  summarise(cities = n()) %>%
  rename(city = City)
```



```{r}
# Get world data frame
country_shapes <- map_data("world") %>%
    rename(country = region) %>%
    rename (city = subregion)

# some basic checks
print("these countries were found in the world_data but not in manybabies data")
setdiff(country_shapes$country,MB.Nlabs$country)
print("these countries were NOT found in the world data but in manybabies data")
setdiff(MB.Nlabs$country,country_shapes$country)

# Merge world_data and world_df
world_data_participating_labs <- MB.Nlabs %>%
  full_join(country_shapes, by = c("country")) %>%
  # change NAs in frequency to 0 (to also plot countries that did not participate)
  mutate(Nlabs = ifelse(is.na(Nlabs), 0,Nlabs))

#Merge world_data and cities_dataframe
cities_dataframe <- MBlabs.Nlabs_cities %>%
  full_join(country_shapes, by = c("city")) %>%
  mutate(MBlabs.Nlabs_cities = ifelse(is.na(city), 0, city))

#renaming cities column to frequency
cities_dataframe <- cities_dataframe %>%
  rename(frequency = cities) 

#keeping only necesary columns
cities_dataframe <- cities_dataframe %>%
  select(city, frequency)

```

# Map data by country and city using ggplot
```{r}

mb1_map <- ggplot() + 
  geom_polygon(data = world_data_participating_labs,  
               aes(x = long, y = lat, group = group, fill = Nlabs)) +                   coord_fixed(1.3) +
  scale_fill_continuous(low="gray85", high="darkred", guide="colorbar",na.value="gray", name = "Labs/country") + 
  geom_point(data = cities_and_lat_log, pch = 21, aes(x = Longitude, y = Latitude, size = n_in_city)) +
  scale_radius(name = "Labs/city", range = c(1,3), breaks = c(1,2,3), labels = c(1,2,3)) +
  ggtitle("Geography of ManyBabies1 Labs") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.text = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

mb1_map

```

#Save plot
```{r}

ggsave("../figures/mb1_map.pdf", mb1_map, device = "pdf", width = 11, height = 6)

```